---
title: "Sweet 16"
date: "3/25/2021"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(rvest)
library(readxl)

Excel_File = "UPDATEDMarch_Madness2021.xlsx"
```

1) Read in Sheets from March_Madness2021 Excel File:

```{r}
Teams_in_Tournament <- read_excel(Excel_File, sheet = "TeamIndex",
                                  col_types = c("numeric", "text", "text"))

TeamStats <- read_excel(Excel_File, sheet="TeamStats")
```

```{r}
Team_For <- read_excel("/Users/Simbo/Desktop/All_TeamStats_2021/Indexed_NCAA_2021_TeamsFor.xlsx")
```

```{r}
Model_df0 <- read_excel("Model_Rev1A.xlsx")

# partition data
set.seed(3.14) #lucky number
train.index0 <- sample(c(1:dim(Model_df0)[1]), dim(Model_df0)[1]*0.6)
train.df0 <- Model_df0[train.index0, ]
valid_df0 <- Model_df0[-train.index0, ]
# -1 suppresses intercept
logit_reg0 <- glm(T1_Wins ~ . -1, 
                  data = train.df0, 
                  family = "binomial")

summary(logit_reg0)
```

Function that returns the team stats used by our model:

```{r}
team_stats <- function(team_index){

    team_2021 <- Team_For %>% 
        filter(TeamIndex == team_index) %>% 
        select("RPG", "SPG", "TOPG", "GP", "FGA", "FG", "X3FGA", "X3FG", "FT") %>% 
        mutate(Avg_FGA = FGA/GP, AVG_3ptA = X3FGA/GP)
    
    return(team_2021)
}
```

Fucntion takes in the two teams in a given matchup and returns the winner.

```{r, echo=TRUE}
predict_winner <- function(team1, team2){
    t1 <- team_stats(team1)
    t2 <- team_stats(team2)
    
    Diff.TRB = t1$RPG - t2$RPG # difference in total rebounds
    Diff.TOV = t1$TOPG - t2$TOPG # difference in turnovers
    Diff.Avg.FGA = t1$Avg_FGA - t2$Avg_FGA
    Diff.STL = t1$SPG - t2$SPG
    Diff.FGp = t1$FG - t2$FG
    Diff.3PA = t1$AVG_3ptA - t2$AVG_3ptA
    Diff.FTp = t1$FT - t2$FT
    Diff.3p = t1$X3FG - t2$X3FG
    
    #Interaction_3 = Diff.3PA * Diff.3p
    #Interaction_FG = Diff.Avg.FGA * Diff.FGp

    df <- c()
    df <- df %>% cbind(Diff.TRB, Diff.TOV, Diff.Avg.FGA, Diff.STL, Diff.FGp, Diff.3PA, Diff.FTp, Diff.3p) %>% as.data.frame()

    pred_winner <- predict(logit_reg0, df, "response")
    
    #print(Model_JMP_Result)
    
    print(pred_winner)
    
    if (pred_winner > 0.5){
        return(team1)
    }else{
        return(team2)
    }
}
```

This function will print the results in a meaningful way.

```{r, echo=TRUE}
print_results <- function(winner, loser){
    win = TeamStats %>% filter(TeamIndex == winner)
    lose = TeamStats %>% filter(TeamIndex == loser)
    
    print(paste0(win$Seed," ", win$School, " (", win$Region, ") beats ", 
                 lose$Seed, " ", lose$School, " (", lose$Region, ").",
                 collapse = "\n"))
    
    return(NULL)
}
```

The following functions take in bracket format and moves winners along.

```{r}
types <- c("numeric", "numeric", "numeric", "text", "numeric", "numeric")

#Excel_File = "/Users/Simbo/Desktop/Spring 2021/Other/minneMUDAC competition/MinneAnalytics/March_Madness2021.xlsx"
#FirstRound <- read_xlsx(Excel_File, sheet = "FirstRound", col_types = types)
#SecondRound <- read_xlsx(Excel_File, sheet = "SecondRound", col_types = types)

#Excel file wth updated teams(actual winners) is the same as the one at the top
Sweet16 <- read_xlsx(Excel_File, sheet = "Sweet16", col_types = types)
Elite8 <- read_xlsx(Excel_File, sheet = "Elite8", col_types = types)
FinalFour <- read_xlsx(Excel_File, sheet = "FinalFour", col_types = types)
Championship <- read_xlsx(Excel_File, sheet = "Championship", col_types = types)
```

This function helps build the model, using the predict_winner() function to update the PredictedWinner column and the team indices for following rounds. It also calls upon the print_result() function to print the predictions.

```{r, echo=TRUE}
bracket_builder <- function(CurrentRound, NextRound){

    for (game in 1:nrow(CurrentRound)){
        winner <- predict_winner(as.list(CurrentRound[game, "Team1Index"]), as.list(CurrentRound[game, "Team2Index"]))
        
        CurrentRound[game, "PredictedWinner"] <- winner
        
        if(CurrentRound[game, "Team1Index"] == winner){
            print_results(winner, as.list(CurrentRound[game, "Team2Index"]))
            
        }else{
            print_results(winner, as.list(CurrentRound[game, "Team1Index"]))
        }
    }
    
    for (i in 1:nrow(NextRound)){
        NextRound[i, "Team1Index"] <- CurrentRound %>% 
                                        filter(NextGameIndex == as.list(NextRound[i, "GameIndex"])) %>% 
                                        slice_head(n=1) %>% select(PredictedWinner)
        
        NextRound[i, "Team2Index"] <- CurrentRound %>% 
                                        filter(NextGameIndex == as.list(NextRound[i, "GameIndex"])) %>% 
                                        slice_tail(n=1) %>% select(PredictedWinner)
    
    }
    
    return(NextRound)
}
```

Now that we have all of the necessary functions, we can use the bracket_builder() function to call each of the rounds, updating the following round with the winning team's indices.

```{r, echo=FALSE}
#print("ROUND OF 64")
#SecondRound <- bracket_builder(FirstRound, SecondRound)

#print("ROUND OF 32")
#Sweet16 <- bracket_builder(SecondRound, Sweet16)

print("SWEET 16")
Elite8 <- bracket_builder(Sweet16, Elite8)

print("ELITE 8")
FinalFour <- bracket_builder(Elite8, FinalFour)

print("FINAL FOUR")
Championship <- bracket_builder(FinalFour, Championship)

print("2021 NCAA CHAMPIONSHIP")
winner <- predict_winner(Championship$Team1Index, Championship$Team2Index)

if(Championship$Team1Index == winner){
    print_results(winner, Championship$Team2Index)
            
}else{
    print_results(winner, Championship$Team1Index)
}
```
