---
title: "Sweet 16"
date: "3/24/2021"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(rvest)
library(readxl)

Excel_File = "UPDATEDMarch_Madness2021.xlsx"
```

1) Read in Sheets from March_Madness2021 Excel File:

```{r}
Teams_in_Tournament <- read_excel(Excel_File, sheet = "TeamIndex",
                                  col_types = c("numeric", "text", "text"))
```

```{r}
TeamStats <- read_excel(Excel_File, sheet = "TeamStats",
                        col_types = c("numeric", "numeric", "text", "text", "text", 
                                      "numeric", "numeric", "numeric", "numeric",
                                      "numeric", "numeric", "numeric", "numeric",
                                      "numeric", "numeric", "numeric", "numeric",
                                      "numeric", "numeric", "numeric", "numeric",
                                      "numeric", "numeric", "numeric"))
```

I don't necessarily think we should use the KenPom data but just in case:

```{r}
kenPom <- read_excel(Excel_File, sheet = "KenPom",
                     col_types = "numeric")
```

Model we used the first time:

```{r}
# This file contained T1_Wins indicator (1 if T1 won, 0 if T2 won), Difference in Win %, 
# and Difference in AdjEm (Strength of Schedule) from the 2015-2020 games for all teams
Model_df0 <- read_excel("test1.xlsx")

# partition data
set.seed(3.14) #lucky number
train.index0 <- sample(c(1:dim(Model_df0)[1]), dim(Model_df0)[1]*0.6)  
train.df0 <- Model_df0[train.index0, ]
valid_df0 <- Model_df0[-train.index0, ]

logit_reg0 <- glm(T1_Wins ~ ., 
                  data = train.df0, 
                  family = "binomial")
```

Function that returns the team stats used by our model:

```{r}
team_stats <- function(team_index){
    # Previous model used win% and AdjEM (strength of schedule) as the variables
    X1 <- TeamStats %>% filter(TeamIndex == team_index) %>% select("W-L%")
    X2 <- kenPom %>% filter(TeamIndex == team_index) %>% select("AdjEM.Strength.of.Schedule")

    team_2021 <- data.frame(X1, X2)
    
    return(team_2021)
}
```

Fucntion takes in the two teams in a given matchup and returns the winner.

```{r, echo=TRUE}
predict_winner <- function(team1, team2){
    t1 <- team_stats(team1)
    t2 <- team_stats(team2)

    dif_WL = t1$W.L. - t2$W.L.
    dif_AdjEm = t1$AdjEM.Strength.of.Schedule - t2$AdjEM.Strength.of.Schedule

    df <- data.frame(dif_WL, dif_AdjEm)
    colnames(df) = c("Dif.Win%", "Dif.AdjEM")
    score <- predict(logit_reg0, df, "response")
    
    if (score >= 0.5){
        return(team1)
    }else{
        return(team2)
    }
}
```

This function will print the results in a meaningful way.

```{r, echo=TRUE}
print_results <- function(winner, loser){
    win = TeamStats %>% filter(TeamIndex == winner)
    lose = TeamStats %>% filter(TeamIndex == loser)
    
    print(paste0(win$Seed," ", win$School, " (", win$Region, ") beats ", 
                 lose$Seed, " ", lose$School, " (", lose$Region, ").",
                 collapse = "\n"))
    
    return(NULL)
}
```

The following functions take in bracket format and moves winners along.

```{r}
types <- c("numeric", "numeric", "numeric", "text", "numeric", "numeric")

# FirstRound <- read_xlsx(Excel_File, sheet = "FirstRound", col_types = types)
# SecondRound <- read_xlsx(Excel_File, sheet = "SecondRound", col_types = types)
Sweet16 <- read_xlsx(Excel_File, sheet = "Sweet16", col_types = types)
Elite8 <- read_xlsx(Excel_File, sheet = "Elite8", col_types = types)
FinalFour <- read_xlsx(Excel_File, sheet = "FinalFour", col_types = types)
Championship <- read_xlsx(Excel_File, sheet = "Championship", col_types = types)
```

This function helps build the model, using the predict_winner() function to update the PredictedWinner column and the team indices for following rounds. It also calls upon the print_result() function to print the predictions.

```{r, echo=TRUE}
bracket_builder <- function(CurrentRound, NextRound){

    for (game in 1:nrow(CurrentRound)){
        winner <- predict_winner(as.list(CurrentRound[game, "Team1Index"]), as.list(CurrentRound[game, "Team2Index"]))
        
        CurrentRound[game, "PredictedWinner"] <- winner
        
        if(CurrentRound[game, "Team1Index"] == winner){
            print_results(winner, as.list(CurrentRound[game, "Team2Index"]))
            
        }else{
            print_results(winner, as.list(CurrentRound[game, "Team1Index"]))
        }
    }
    
    for (i in 1:nrow(NextRound)){
        NextRound[i, "Team1Index"] <- CurrentRound %>% 
                                        filter(NextGameIndex == as.list(NextRound[i, "GameIndex"])) %>% 
                                        slice_head(n=1) %>% select(PredictedWinner)
        
        NextRound[i, "Team2Index"] <- CurrentRound %>% 
                                        filter(NextGameIndex == as.list(NextRound[i, "GameIndex"])) %>% 
                                        slice_tail(n=1) %>% select(PredictedWinner)
    
    }
    
    return(NextRound)
}
```

Now that we have all of the necessary functions, we can use the bracket_builder() function to call each of the rounds, updating the following round with the winning team's indices.

```{r, echo=FALSE}
# print("ROUND OF 64")
# SecondRound <- bracket_builder(FirstRound, SecondRound)

# print("ROUND OF 32")
# Sweet16 <- bracket_builder(SecondRound, Sweet16)

print("SWEET 16")
Elite8 <- bracket_builder(Sweet16, Elite8)

print("ELITE 8")
FinalFour <- bracket_builder(Elite8, FinalFour)

print("FINAL FOUR")
Championship <- bracket_builder(FinalFour, Championship)

print("2021 NCAA CHAMPIONSHIP")
winner <- predict_winner(Championship$Team1Index, Championship$Team2Index)

if(Championship$Team1Index == winner){
    print_results(winner, Championship$Team2Index)
            
}else{
    print_results(winner, Championship$Team1Index)
}
```

